---
title: "Introduction to exchanger"
author: "Neil Marchant"
date: "17 December 2020"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to exchanger}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=7, 
  fig.height=5
)
set.seed(319158)
```

We demonstrate how to perform entity resolution using the `RLdata500` test 
data set included with the package. `RLdata500` contains 500 synthetic 
personal records, 50 of which are duplicates with randomly-generated errors.
In the code block, below we load the package and examine the first few rows 
of `RLdata500`. 
```{r, eval=TRUE, message=FALSE, warning=FALSE}
library(exchanger)
library(comparator)
library(clevr)
head(RLdata500)
```

Next we specify the model parameters for the entity attributes. 
For simplicity, we use a uniform prior for the distortion probability 
associated with each attribute.
```{r, eval=TRUE, message=FALSE, warning=FALSE}
unif_prior <- BetaRV(1, 1)
```

We model the distortion for the name attributes (`fname_c1` and `lname_c1`) 
using Levenshtein (edit) distance 
The `transform_dist_fn` applies a threshold at a distance of 3, so that any 
pair of names `y`, `x` with an edit distance strictly greater than 3 are 
assumed to be "infinitely" apart. 
This improves computational efficiency, as the inference algorithm exploits 
the fact that `x` can never be a distortion of `y` if their distance is 
strictly greater than 3.
For the date attributes (`by`, `bm` and `bd`) we assume a constant distance 
function, which means all values in the domain are equally likely a priori 
as a distorted alternative. 
The `CategoricalAttribute` constructor provides a shortcut for an attribute 
with a constant distance function.
```{r, eval=TRUE, message=FALSE}
attr_params <- list(
  "fname_c1" = Attribute(transform_dist_fn(Levenshtein(), threshold = 3.0),
                         distort_prob_prior = unif_prior),
  "lname_c1" = Attribute(transform_dist_fn(Levenshtein(), threshold = 3.0),
                         distort_prob_prior = unif_prior),
  "by" = CategoricalAttribute(distort_prob_prior = unif_prior),
  "bm" = CategoricalAttribute(distort_prob_prior = unif_prior),
  "bd" = CategoricalAttribute(distort_prob_prior = unif_prior)
)
```

Finally we specify the prior over the linkage structure (clustering). Here we 
use a Pitman-Yor random partition for the prior, with hyperpriors on the 
concentration and discount parameters.
```{r, eval=TRUE, message=FALSE}
clust_prior <- PitmanYorRP(alpha = GammaRV(1, 1), d = BetaRV(1, 1))
```

All that remains is to initialize the model and run inference. 
For demonstration purposes, we only generate 100 posterior samples.
```{r, eval=TRUE, message=FALSE}
model <- exchanger(RLdata500, attr_params, clust_prior)
system.time(
result <- run_inference(model, n_samples=100, thin_interval=10, burnin_interval=1000)
)
```

We recommend inspecting trace plots to verify that the Markov chain has 
reached equilibrium and is mixing well. The results below seem acceptable given 
the small number of samples.
```{r}
n_linked_ents <- extract(result, "n_linked_ents")
distort_probs <- extract(result, "distort_probs")
plot(n_linked_ents)
plot(distort_probs)
```

We can obtain a point estimate of the most likely linkage structure using the 
_shared most probable maximal matching sets method_.
```{r, eval=TRUE}
pred_clusters <- smp_clusters(result)
```

Since RLdata500 comes with ground truth, we can evaluate the predicted 
linkage structure using supervised metrics. We find that the model performs 
well on this data set, achieving high pairwise precision and recall.
```{r, eval=TRUE}
n_records <- nrow(RLdata500)
true_pairs <- membership_to_pairs(identity.RLdata500)
pred_pairs <- clusters_to_pairs(pred_clusters)
measures <- eval_report_pairs(true_pairs, pred_pairs, num_pairs=n_records*(n_records-1)/2)
print(measures)
```

